### 2018-06-04 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\ection{Алгоритми со длабоко учење}

@@

Првиот документиран обид за користење на невронски мрежи за генерирање на музика е од страна на Д. Ек во трудот \cite{Eck2002}. Целта е тренирање на модел кој ќе генерира блуз во 12 такта. Музиката се состои од 2 оделни компоненти, секвенца на акорди кои го одредуваат ритамот и ја даваат основната структура на музиката и мелодиска линија. Моделот има две групи од по 4 пара LSTM (Long Short Term Memory) ќелии (рекурентни неврони), од кои едната е задолжена за учење на мелодијата, а друга за акордната секвенца. Музиката ја преставуваат во тн. репрезентација во пиано лента, која е дво димензионална бинарна матрица во која едната оска (подолгата) го претставува времето, а другата оска го преставува множеството од можни тонови. Музиката има строга структура, 12 такта, 4/4 ритам и времето е подделено на 1/8 ноти. Моделот го испробале на два експерименти, во првиот тој ја учи само секвенцата на акорди, а во вториот и акордите и мелодијата. Тренирањето е извршено со cross-entropy фунција на загува, со сигмоидална функција на активација. Делот за учење на акорди ја споделува својата внатрешна состојба со делот за учење на мелодијата. Авторот смета дека системот генерира добра музика, но сепак проблемот што пробува да го реши е хипер фокусиран и ограничен. Трудот \cite{Eck2008} претставува надоградување на претходно опишаниот систем. Го нема веќе фокусот на блуз во 12 такти, туку работат со податочно множество од поразлини MIDI датотеки, зголемена е комплексноста на моделот, многу повеќе ќелии, повеќе слоеви од истите. Додатно

Long-timescale structure is fundamental to music, so all architectures try to capture it. However regardles of the architecture the possibilities explode with the timescale. Their solution primarily deals with sequences extracted from MIDI, not performed music. The core algorithms are expected to be well suited for digital audio as well. Improvement over the previous work is addition of time-delay connections that correspond to the metrical hierarchy of a particular piece. This meter information should provide the LSTM net with musically important temporal structure. With it the model was able to capture some of the repetitive structure crucial to learning a musical style. Meter is the sense of strong and weak beats that arises from the interaction among hierarchical levels of sequences having nested periodic components. Such a hierarchy is implied in Westernmusic notation, where different levels are indicated by kinds of notes (whole notes, half notes, quarter notes, etc.) and where bars establish measures of an equal number of beats (Handel, 1993). For instance, most contemporary pop songs are built on four-beat meters. In such songs, the first and third beats are usually emphasized. Knowing the meter of a piece of music helps in predicting other components of musical structure such as the location of chord changes and repetition boundaries (Cooper and Meyer, 1960). Meter provides us with key information about musical structure. Music, at least popular Western music, tends to be chunked in ways that correspond to meter. Chord changes, for example, usually occur on metrical boundaries. Also, music tends to repeat at intervals corresponding to the metrical hierarchy. A dynamic learner such as a recurrent neural network (details described below) could learn to repeat a fixed number of learned patterns, but it could not learn to repeat an arbitrary sequence because it has no way to implement content-addressable memory. Meter is provided to the network in the form of delayed inputs. If music is sampled k times a measure, eg k=8, for 4/4 meter to make it easy for the network to correlate metrically-relevant delays by providing time-delayed copises of the input at times t -15, t-31 and t-63 (2,4 and 8 measures). The recurrent network will learn the regular sequence on its own, but lags related to meter are give more attention. Theiir LSTM network has a standard feed-forward layer in paralel to the LSTM cells. It should accelerate training by learning local dependencies quicker than LSTM alone, leaving the LSTM cells to handle longer timescale dependencies. This reulted in melodies with smoother contours. The network is trained as a next-step predictor. It trains to predict probability density over all possible notes at time t using as input the note (and chord) values at time t-1. In this model the network recieves as input not just the sequence delayed by a single lag (t-1) but also delayed by lags correpsonding to metrical structure of the piece. Multiple songs are presented to the network in a single sequence. Error flow is truncated at song boundaries, so that the network doesn't learn spurious corelations from one song to another. With the trained model, music can be generated by presenting it with the first few notes of a song that it has never seen in training and using the network predictions to generate network inputs. Network predictions are conditioned using a softmax function, ensuring that the sum of the output vector is 1.0. This allows us to interpret the output vector as a probability estimation from which we can select the next note. The selected note is then presented to the network at the next timestep as an input. For some simulations, we applied a threshold to our note generation, ensuring that very low probability notes would not be chosen. The threshold we used was 1/N where N is the cardinality of the output vector. MIDI data representation, turned into input vectors by sampling the file at eight-note intervals. Most other data formats can be turned into MIDI files. Number of octaves limited between C3 and C5, notes outside of range folded into range. Time is represented implicitly, so no note durations are present in i/o. Identical to previoys works. Notes are represented locally in a 1-hot vector. Notes not in the corpus are not represented. Chords are also encoded locally in 1-hot vector. So eg. Fmaj7 is represented by only 1 input element and not as several units representing the notes that make up the chord. Advantage of this aproach is the network can more quickly learn chord structure, but distadvantage is the network cannot generalize to unseen chords. Several 1-hot vectors are added to encode metrical structure, corressponding to the time-delayed versions of the input. Dataset: They used Irish reels in 4/4. They selected reels from thesession.org that were in C maj and E maj. It's a good dataset with at least 1700 reels. It however does not have consistant chord information on all songs, so they used only the melodies. They also used the Nottingham dataset, transposing all the songs in the same key. It does provide good chord information that they used. THe most interesting experiment, run on Nottingham datase, contains both melody and chords. Take care not to overfit the dataset with too hight a network capacity (too many nodes) or too much training, it'll just repeat the same notes. When training on melodies and chrods, they achieved a result where the melody does follow the chords. Also the model did capture the slow-changing chord structures and is able to synchronize faster-changing melodic structure with the chords.

>>>>

\section{Алгоритми со длабоко учење}

@@

Првиот документиран обид за користење на невронски мрежи за генерирање на музика е од страна на Д. Ек во трудот \cite{Eck2002}. Целта е тренирање на модел кој ќе генерира блуз во 12 такта. Музиката се состои од 2 оделни компоненти, секвенца на акорди кои го одредуваат ритамот и ја даваат основната структура на музиката и мелодиска линија. Моделот има две групи од по 4 пара LSTM (Long Short Term Memory) ќелии (рекурентни неврони), од кои едната е задолжена за учење на мелодијата, а друга за акордната секвенца. Музиката ја преставуваат во тн. репрезентација во пиано лента, која е дво димензионална бинарна матрица во која едната оска (подолгата) го претставува времето, а другата оска го преставува множеството од можни тонови. Музиката има строга структура, 12 такта, 4/4 ритам и времето е подделено на 1/8 ноти. Моделот го испробале на два експерименти, во првиот тој ја учи само секвенцата на акорди, а во вториот и акордите и мелодијата. Тренирањето е извршено со cross-entropy фунција на загува, со сигмоидална функција на активација. Делот за учење на акорди ја споделува својата внатрешна состојба со делот за учење на мелодијата. Авторот смета дека системот генерира добра музика, но сепак проблемот што пробува да го реши е хипер фокусиран и ограничен. Трудот \cite{Eck2008} претставува надоградување на претходно опишаниот систем. Го нема веќе фокусот на блуз во 12 такти, туку работат со податочно множество од поразлини MIDI датотеки, зголемена е комплексноста на моделот, со многу повеќе ќелии и повеќе слоеви од истите. Додатно на моделот му прикажуваат и временски дилатирани копии од податоците, на семантички важни растојание. Ова доаѓа од хипотезата дека во музиката значајни настани се случуваат на метрички важни места како што се почетокот и крајот на тактовите, како и на моменти зависни од ритамо (пример во македонскиот 7/8 ритам ен-два, ен-два, ен-два-три, има значајни транзиции на секое „ен“). Ова го постигнуваат со проширување на пиано лентата да при секој чекор на моделот му се прикаже лентата од тековниата музика, и од музиката од претходно зададени интервали, како на пр. пред 1, 4 и 16 такта толно на истиот удар во тактовите. Покрај LSTM слоеви моделот има и паралелен целосно поврзан слој неврони којшто треба да помогне во учење на локални зависности меѓу нотите. Мрежата ги учи песните како една секвенца, со што грешката при учење се ресетира акумулираната грешка на граница меѓу песните. Нотите се ограничени помеѓу C3 и C5 и сите песни се транспонирани во ист клуч. Користени се ирски народни песни во 4/4 ритам. Моделот генерира нови песни со предвидување на следната нота откако ќе се припреми претходно со исечок од постоечка песна.
